---
title: "Essentials Week 05: File Paths & Tables"
author: "DE"
date: "03/03/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T)
```

```{r pckgs, include=F}

# task 1: load packages tidyverse, knitr, kableExtra & here





# task 2: run this chunk & look at the output in the console to see where here() starts

```

```{r data-1, include=F}

# read in data_1 here

data_1 <- 

```

```{r data-2, include=F}

# read in data_2 here

data_2 <- 

```


<!-- I'll use this Rmd doc to demonstrate different functions & arguments using data_1, then you can try it out yourself with data_2 -->


# knitr::kable() arguments  

*knitr::kable() is used to generate tables. There are a number of arguments to the kable() function, the most useful/common being col.names, caption, & digits*    

<br>

```{r table-1}

knitr::kable(data_1)

# this is the original data in a table with no additional formatting

```

<br>

```{r col-names}

knitr::kable(data_1, col.names = c("Track Name", "Artists", "Danceability", "Energy", "Liveness", "Tempo", "Valence", "Popularity", "Track Duration"))

# note that the number of column names you give in this argument MUST match the number of columns in the dataset

```

<br>

```{r caption}

knitr::kable(data_1, col.names = c("Track Name", "Artists", "Danceability", "Energy", "Liveness", "Tempo", "Valence", "Popularity", "Track Duration"), caption = "This is a pretty table now it has a caption.")

# remember that the caption has to be enclose within speech marks

```

<br>

```{r digits}

knitr::kable(data_1, col.names = c("Track Name", "Artists", "Danceability", "Energy", "Liveness", "Tempo", "Valence", "Popularity", "Track Duration"), caption = "This is a pretty table now it has a caption.", digits = 1)

# here im specifying the number of digits for all columns in my table

```

<br>

# Your turn to paRty part 1!

## Tasks

Using the data_2 dataset, create a table with the below specification    

- change the column names to be: Track Name, Artists, Energy, Popularity, Tempo, Valence, Track Duration  
- add a caption to describe your table  
- show values to 2 decimal places  

<br>

```{r data-2-table}

# create your table here





```

<br>

# kableExtra::kable_styling() arguments  

*kableExtra::kable_styling() is not for generating tables, instead, it adds features to tables with the pipe %>%.  There are a number of options you can set to add/change the formatting of the __whole__ table, the key ones are font_size, position, & bootstrap_options, that will make your table nice and pretty*  

<br>

```{r font-size}

knitr::kable(data_1) %>%
  kableExtra::kable_styling(font_size = 8)

# this sets the font size

```

<br>

```{r width}

knitr::kable(data_1) %>%
  kableExtra::kable_styling(font_size = 8, full_width = FALSE)

# here we're setting the font size, whether the table should be the full width (which ive set to FALSE)

```

<br>

```{r position}

knitr::kable(data_1) %>%
  kableExtra::kable_styling(font_size = 8, full_width = FALSE, position = "left")

# here we're setting the font size, whether the table should be the full width (which ive set to FALSE) and the position of the table

```

<br>

```{r bootstrap-styling}

knitr::kable(data_1) %>%
  kableExtra::kable_styling(font_size = 8, position = "center", bootstrap_options = "striped")

# here we're setting the font size, table positioning & appearance to be striped

```

<br>

```{r bootstrap-styling-2}

knitr::kable(data_1) %>%
  kableExtra::kable_styling(font_size = 8, position = "center", bootstrap_options = "hover")

# here we're setting the font size, table positioning & appearance to change colour when we hover over a row

```

<br>

# Combining it all together

*We can easily use kable() and kable_styling() arguments in one pipe to make the most beautiful tables we ever did see*

<br>

```{r table-combo}

knitr::kable(data_1, col.names = c("Track Name", "Artists", "Danceability", "Energy", "Liveness", "Tempo", "Valence", "Popularity", "Track Duration"), caption = "This is a pretty table now it has a caption.", digits = 1) %>%
  kableExtra::kable_styling(font_size = 8, position = "center", bootstrap_options = "hover")

# here i'm just specifying all our kable() arguments that we used before, then using a pipe to our kableExtra arguments

```

<br>

# Your turn to paRty part 2!  

## Tasks  

1. Recreate your data_2 table in the chunk below with kable() as you did in the first task above  

2. Style your table with kable_styling so that the font size is 12, the full width is FALSE, the position is right, and the appearance (bootstrap_options) is "bordered"  

<br>

```{r data-2-table-styling}

# style your table here by specifying the font size, the width, the positioning, and the bootstrap options





```

<br>





